---
title: "Application Qian 2001 method on Selevsek 2015 data (Z-normalize by genes)"
author: "s_morimoto"
date: "2016/07/01"
output: 
  pdf_document: 
    latex_engine: xelatex
---
This source code is built under R ver.3.1.2  

You are using  `r R.version.string`  

The transcriptome data are taken in *Lee,2011*, described in *Selevsek,2015*.
The data has 6 time points(T1~T6), but T6(240[min]) is twice of T5(120[min])
so discarded from analysis.  

The proteome data has 6 time points.
This data has T2=15[min] in which timepoint was not taken data in *Lee,2011*.
To establish consistency to transcriptome data, deleted T2(15[min])data, and
rename T3 to T2, T4 to T3, and so on.

SETUP parameters
```{r setting up parameters}
rm()

dataDirectry   <- "/Users/mos/Dropbox/Draft_201603Morimoto/Analysis/Data"

dataPath_trc     <- "Gene2011_testResult_final.xlsx" # 
dataPath_prt     <- "ProtSWATH_testResult_final.xlsx" #  

sheet = 1
header=TRUE

## filenames of the OUTPUT FILEs

outputDirectry <- "/Users/mos/Dropbox/Draft_201603Morimoto/Analysis/Data"

dataOutput_trc.prefix     <- "Gene2011_testResult_ana_20160706" # 
dataOutput_prt.prefix     <- "ProtSWATH_testResult_ana_20160706" #  

file_dataGQ         <- 'dataGQ_20160706.csv'

file_save.image     <- 'Selevsek_Qian_20160706.RData'

```

LOAD PACKAGES

```{r install packages,echo=FALSE}

## avoid the error (source: https://code.google.com/p/rexcel/issues/detail?id=33)
#options(java.parameters = "-Xmx1000m") 

#install.packages('xlsx')
#install.packages('readxl')
#install.packages('data.table')
#install.packages('kohonen')
#install.packages('plyr')
#install.packages('dplyr')
#install.packages('tidyr')
#install.packages('pander')
#install.packages('pspearman')
#install.packages('SuppDists')
#install.packages('Kendall')
#install.packages('stringr')

require('plyr')
require('dplyr')
require('tidyr')
require('xlsx')
require('readxl')
require('ggplot2')
require('reshape2')
require('pander')
require('stringr')

panderOptions('table.split.table', 1000)
panderOptions('table.continues', '')

```



LOAD DATA
```{r loading data }

  
raw_data.trc <- read_excel(
  sprintf(fmt = '%s/%s',
          dataDirectry ,
          dataPath_trc
          ),
  sheet,
  col_names = header,
  col_types = NULL, 
  na = "",skip = 0
  )[ ,2:4] 

raw_data.prt <- read_excel(
  sprintf(fmt = '%s/%s',
          dataDirectry ,
          dataPath_prt
          ),
  sheet,
  col_names = header,
  col_types = NULL,
  na = "",skip = 0
  )[ ,2:4] 

data.trc <- data.frame(Protein=unique(raw_data.trc$Protein),
                       log2FC=0,
                       Label='T1-T1'
                       ) %>%
  full_join(raw_data.trc) %>%
  mutate(
         id=as.character(Protein),
         var=as.character(Label),
         val=as.numeric(log2FC),
         dtname='trc'
         ) %>%
  mutate(
    var = 
      ifelse(Label=='T1-T1','0min',
             ifelse(Label=='T2-T1','30min',
                    ifelse(Label=='T3-T1','60min',
                           ifelse(Label=='T4-T1','90min',
                                  ifelse(Label=='T5-T1','120min',
                                         ifelse(Label=='T6-T1','240min',NA)
                                         )
                                  )
                           )
                    )
             )
    ) %>%
  dplyr::select(id,var,val,dtname)

data.prt <- data.frame(Protein=unique(raw_data.prt$Protein),
                       log2FC=0,
                       Label='T1-T1'
                       ) %>%
  full_join(raw_data.prt) %>%
  mutate(
         id=as.character(Protein),
         val=as.numeric(log2FC),
         dtname='prt'
         ) %>%
  mutate(
    var = 
      ifelse(Label=='T1-T1','0min',
             ifelse(Label=='T2-T1','15min',
                    ifelse(Label=='T3-T1','30min',
                           ifelse(Label=='T4-T1','60min',
                                  ifelse(Label=='T5-T1','90min',
                                         ifelse(Label=='T6-T1','120min',NA)
                                         )
                                  )
                           )
                    )
             )
    ) %>%
  dplyr::select(id,var,val,dtname)
```

```{r T1 is 1}
# This chank is added 2016/7/1

data.trc <- data.trc %>%
  mutate(val = 2^val)

data.prt <- data.prt %>%
  mutate(val = 2^val)

```


```{r}
write.xlsx(
  data.trc,
  sprintf(fmt = '%s/%s.xlsx',
  outputDirectry,
  dataOutput_trc.prefix
  )
  )

write.xlsx(
  data.prt,
  sprintf(fmt = '%s/%s.xlsx',
  outputDirectry,
  dataOutput_prt.prefix
  )
  )

```
