---
title: "Application Qian 2001 method on Fournier2015 data (Z-normalized by genes)"
author: "s_morimoto"
date: "2015/11/20"
output: pdf_document
---
This source code is built under R ver.3.1.2

You are using  `r R.version.string` 



SETUP parameters
```{r setting up parameters}
rm()

##--- setting : data  ##
dataDirectry   <- "/Users/mos/Dropbox/D/Thesis001/Output"

dataPath_trc     <- "Gene2011_testResult_ana.xlsx"
                    #'Fournier_20151018_trc.xlsx'
                    #"Gene2011_testResult_ana.xlsx" # 
dataPath_prt     <- "ProtSWATH_testResult_ana.xlsx"
                    #'Fournier_20151018_prt.xlsx'
                    #"ProtSWATH_testResult_ana.xlsx" # 
dataPath_pathway <- "2015_Selevsek_proteins_KEGG.xlsx" #  
bioproc <- 'GlySerThrMetabo'

typeBoolean  <- FALSE
sheet    = 1
sheetPathway = 1

header   = TRUE

startRow_trc = 1
startRow_prt = 1
startRow_pw  = 1

endRow_trc = 40027 #2296
endRow_prt = 15535 #2296
endRow_pw  = 54#54

timePoint <- 5
repet     <- 1

discrete_x_label <- c('inter-temporal')
continuous_y_label <- c('log2 fold change')


cols <- c('T1-T1','T2-T1','T3-T1','T4-T1','T5-T1')
        #c('min0','min20','min40','min60','min120','min240','min360')
        #c('T1-T1','T2-T1','T3-T1','T4-T1','T5-T1')

## end of setting : data  ---##

columnName <- c(rep(cols,repet))

## --- setting : data processing before analysis #

z_norm <- TRUE  ## Z-Normalize by genes and datanames ( TRUE/ FALSE)
interTemporal <- TRUE

## end of setting : data processing before analysis ---##

## --- setting : filenames of the OUTPUT FILEs ##

allOutput.prefix <- 'Selevsek_2015'

outputDirectry <- "/Users/mos/Dropbox/D/Thesis001/Output"

file_dataGQ.prefix         <- 'dataGQ_Selevsek_20151125'
file_data_idConvert.prefix <- 'id.convert'
file_save.image     <- 'Selevsek_20151125.RData'

## end of setting : filenames of the OUTPUT FILEs ---##

```

```{r BioMart setting}

 # Gene id convert (via bioMart)
# 
#bioMartDB      <- "fungi_mart_29" # 2015.10.28 DB update                                     # biomaRt::listMarts()
#bioMartDataSet <- 'scerevisiae_eg_gene'                     
#                                  # biomaRt::listDatasets(db)
#inputName      <- 'wikigene_name'   # 'wikigene_name' or 'ensembl_gene_id' 
#                                  # biomaRt::listFilters(sceg)
#outputName     <- 'uniprot_swissprot_accession'
#                                  # biomaRt::listFilters(sceg)

bioMartHost    <- 'fungi.ensembl.org' # 2015.11.10 updated (HOST='biomart.org' has stopped)
bioMartDB      <- "fungal_mart" # 2015.11.10 DB update    
                                  # biomaRt::listMarts(host=bioMartHost)
bioMartDataSet <- 'scerevisiae_eg_gene'                     
                                  # biomaRt::listDatasets(db)
inputName_1      <-  'ensembl_gene_id' # 'wikigene_name' or 'ensembl_gene_id' 
                                  # biomaRt::listFilters(sceg)
outputName     <- 'uniprot_swissprot_accession'
                                  # biomaRt::listFilters(sceg)s

```

```{r functions sorce code file}

funcDirectry <- '/Users/mos/Dropbox/D/Thesis001/PG/Functions/'

EscoreCalc <- 'func_for_calcEscore_20151124.R'

```


LOAD PACKAGES

```{r install packages,echo=FALSE}

## avoid the error (source: https://code.google.com/p/rexcel/issues/detail?id=33)
#options(java.parameters = "-Xmx1000m") 


## LIBRARIES
rm()

packages <- c(
  'dplyr', # progress bar
  'plyr',  # progress bar
  'dplyr', # progress bar
  'tidyr',
  'xlsx',
  'readxl',
  'ggplot2',
  'gplots',
  'GMD',
  'pvclust',
  'reshape2',
  'pander',
  'stringr'
#  'biomaRt'
  ) 

new.packages <-
  packages[!(packages %in% installed.packages())] 
  # installed.packages() returns installed packages 

if(length(new.packages) > 0){ 
  install.packages(new.packages, repos='http://cran.us.r-project.org')
}
require('plyr')  # progress bar
require('dplyr') # progress bar
require('tidyr')
require('xlsx')
require('readxl')
require('ggplot2')
require('gplots')
require('GMD')
require('pvclust')
require('reshape2')
require('pander')
require('stringr')

if( !("biomaRt" %in% installed.packages()) ){
  source('http://bioconductor.org/biocLite.R')
  biocLite( 'biomaRt' )
  require( 'biomaRt' )
}else  require( 'biomaRt' )

#```

panderOptions('table.split.table', 1000)
panderOptions('table.continues', '')

```

definition of functions

```{r defFUN makeDifData}
makeDifData <- function(data,var){
  for (i in 2:length(var)){
    end   <- var[i]
    start <- var[i-1]
    rescol <- paste('d',end,sep='_')
    ddata  <- data[,end] - data[,start]
    data[,rescol] <- ddata
  }
  return(data)
  }
```


Qian,Gerstein (J.Mol.Biol.,2001)
```{r definition of FUNC gq_method}

source(file = sprintf(fmt = '%s%s',funcDirectry,EscoreCalc))

##--- Example : function gq_method ##

gq_method(c(0,0,0,0,1,1,1,-1,1,1,1,1),c(0,0,0,0,1,1,1,1,1,1,1,-1),4,3,'pos')

gq_method(c(-1.597670,1.722744,1.699152,-2.219394,4.418399,2.081298,4.085573),
          c(-5.823882,9.529135,-5.602938,2.634425,3.772645,-8.844689,-3.819055),7,1,'neg')

## end of Example : function gq_method ---##
```


LOAD DATA
```{r loading data }

raw_data.trc <- read_excel(
  sprintf(fmt = '%s/%s',
          dataDirectry,
          dataPath_trc
          ),
  sheet,
  col_names = header,
  col_types = NULL, 
  na = "",
  skip = startRow_trc-1
  )
raw_data.prt <- read_excel(
  sprintf( fmt = '%s/%s',
           dataDirectry,
           dataPath_prt), 
  sheet,
  col_names = header,
  col_types = NULL, 
  na = "",
  skip = startRow_trc-1
  )
```

```{r tranforming data}
data.trc <- mutate(raw_data.trc,
                   dtname='trc'
                   ) %>%
  dplyr::select(id,var,val,dtname)

data.prt <- mutate(raw_data.prt,
                   dtname='prt'
                   ) %>%
  dplyr::select(id,var,val,dtname)


dataWide.trc <- data.trc %>%
  spread(key=var,value=val)
dataWide.prt <- data.prt %>%
  spread(key=var,value=val)

if (z_norm==TRUE){
  dataLong <- bind_rows(data.trc,data.prt) %>%
    rename(val2=val) %>%
    group_by(id,dtname) %>%  # " dataLong <- as.data.frame(dataLong) "
    mutate(val=scale(val2, center = TRUE, scale = TRUE)) %>%
    dplyr::select(-val2)
}else 
  dataLong <- bind_rows(data.trc,data.prt) 

attributes(dataLong$val) <- NULL 
  # attrs are created by scale function which causes errors when this data treated as data.frame
dataLong <- as.data.frame(dataLong) 
  # ungroup the BY-groups created by  " %>% group_by(id,dtname)) "

```


LOAD gene filtering DATA
```{r filtering data }

data_ana <- 
  dataLong

#data_ana$var <- factor(data_ana$var,levels=cols)

if(1-is.null(bioproc)){
  pathwayData   <- read.xlsx(
    sprintf(fmt = '%s/%s',
            dataDirectry,
            dataPath_pathway),
    sheetPathway, 
    header=header,
    startRow=startRow_pw, 
    endRow=endRow_pw,
    colIndex=1:3, dtname='pw') %>%
    filter(pathway==bioproc) %>%
    mutate(id=as.character(id),
           dtname=as.character(dtname),
           protein=as.character(protein)
           )
  data_ana <- inner_join(dataLong,
                         pathwayData %>% dplyr::select(-dtname)
                         ,by='id')
  summ_data_ana <- data_ana %>%
    group_by(dtname,pathway,var) %>%
    summarise(
      n=n(),mean=mean(val),sd=sd(val),
      min=min(val),median=median(val),max=max(val))
  pander(summ_data_ana)
  }else{
    summ_data_ana <- data_ana %>%
      group_by(dtname,var) %>%
      summarise(
        n=n(),mean=mean(val),sd=sd(val),
        min=min(val),median=median(val),max=max(val))
    pander(summ_data_ana)
    }

```



```{r making time dif data}


w.timePoint <- timePoint

if(interTemporal==TRUE){
  difdata_ana <- data_ana
  difdata_ana <- makeDifData(difdata_ana%>%spread(key=var,value=val),cols) %>%
    dplyr::select(id,dtname,starts_with('d_')
           ) %>%
    gather(var,val,starts_with('d_'))
  data_ana <- difdata_ana
  timePoint <- timePoint-1
}
```


```{r}

if(interTemporal==TRUE){
  dataGQ <- data_ana %>% 
    spread(key=var,value=val) %>%
    dplyr::select(id,dtname,starts_with("d_"))
  }else{
  dataGQ <- data_ana %>% 
    spread(key=var,value=val) %>%
    dplyr::select(id,dtname,one_of(cols))
  }

write.csv(dataGQ,
          sprintf(fmt = '%s/%s_%s_output.csv',
                  outputDirectry,
                  allOutput.prefix,
                  file_dataGQ.prefix
                  )
          )

```

```{r using BioMart}

db.DL <- useMart(bioMartDB,host=bioMartHost) # listMarts(host=bioMartHost)
sceg <- useDataset(bioMartDataSet, mart = db.DL) # listDatasets(db)

id_convert <- getBM(
  attributes = c(inputName_1,outputName),
  filters =  c(inputName_1), # listFilters(sceg)
  values = dataGQ$id, #
  mart = sceg
  )
id_convert[,'id']    <- id_convert[,inputName_1]
id_convert[,'uniid'] <- id_convert[,outputName]

if(is.null(bioproc)) bioproc <- 'all'
write.csv(id_convert %>%
            dplyr::select(id,uniid),
          sprintf(fmt = '%s/%s_%s_%s.csv',
                  outputDirectry,
                  allOutput.prefix,
                  file_data_idConvert.prefix,
                  bioproc)
          )
```

Qian,Gerstein (J.Mol.Biol.,2001)
```{r }
#gq_list_pos <- list()
#for(i in 1:length(unique(dataGQ$id))){
#  gene <- unique(dataGQ$id)[i]
#  data <- dataGQ %>%
#    filter(id==gene)
#    if(nrow(data)>=2){
#      gq_res <-  gq_method(data[1,3:length(data)],
#                           data[2,3:length(data)],
#                           timepoint=timePoint,
#                           rep=repet,'pos')
#      gq_list_i <- list()
#      gq_list_i[[1]] <- data
#      gq_list_i[[2]] <- gq_res
#      gq_list_pos[[i]] <- gq_list_i
#      }
#  }
#pander(gq_list_pos)
```

```{r saving RData file}
if(is.null(bioproc)) bioproc <- 'all'
  
save(
  dataGQ,
  timePoint,
  repet,
  bioproc,
  discrete_x_label,
  continuous_y_label,
  cols,
  gq_method,
  file = sprintf(fmt = '%s/%s_%s_%s',
                   outputDirectry,
                   allOutput.prefix,
                   bioproc,
                   file_save.image
                   )
     )

